Factors affecting health sector expenditure in OECD countries

### 2025-05-03 code/analysis-code/statistical-analysis.qmd
made tables formal

modelsummary(
  model,
  output = here("results", "tables", "primary_regression_results.png"),
  title = "This table represents the primary regresssion model",
  stars = c('*' = 0.1, '**' = 0.05, '***' = 0.01),
  notes = list(
    "Standard errors are shown in parentheses.",
    "Significance levels: * p < 0.1, ** p < 0.05, *** p < 0.01."
  )
)

@@

# Create a data frame with the results
dw_results <- data.frame(
  Statistic = dw_statistic,
  P_Value = dw_p_value
)

# Generate a table using gt
dw_table <- gt(dw_results) %>%
  tab_header(
    title = "Durbin-Watson Test Results"
  ) %>%
  fmt_number(
    columns = vars(Statistic, P_Value),
    decimals = 4
  )

>>>>

# Define a named vector for clean variable names
clean_names <- c(
  "(Intercept)" = "Intercept",
  "GDP_growth_annual_" = "GDP Growth (annual %)",
  "Income_share_held_by_lowest_20" = "Income Share (Lowest 20%)",
  "Control_of_Corruption_Estimate" = "Control of Corruption",
  "Life_expectancy_at_birth_total_years" = "Life Expectancy (years)",
  "Unemployment_youth_total__of_total_labor_force_ages_1524_modeled_ILO_estimate" = "Youth Unemployment (% of labor force, ages 15-24)",
  "Trade__of_GDP" = "Trade (% of GDP)",
  "Foreign_direct_investment_net_inflows_BoP_current_US" = "FDI Net Inflows (current US$)",
  "Access_to_electricity_rural__of_rural_population" = "Rural Electricity Access (% of rural population)",
  "Age_dependency_ratio__of_workingage_population" = "Age Dependency Ratio"
)

# Generate the table with formatted names
modelsummary(
  model,
  output = here("results", "tables", "primary_regression_results.png"),
  title = "",
  stars = c('*' = 0.1, '**' = 0.05, '***' = 0.01),
  coef_rename = clean_names,
  notes = list(
    "Standard errors are shown in parentheses.",
    "Significance levels: * p < 0.1, ** p < 0.05, *** p < 0.01."
  )
)

@@

# Create a data frame with the results
dw_results <- data.frame(
  Statistic = dw_statistic,
  `p-value` = dw_p_value,  
  check.names = FALSE      # Preserve the column name formatting
)

# Generate a table using gt
dw_table <- gt(dw_results) %>%
  tab_header(
    title = "Durbin-Watson Test Results"
  ) %>%
  fmt_number(
    columns = everything(),  # Applies to all columns
    decimals = 4
  )

### 2025-05-03 code/robustness-code/robustness-code.qmd
made tables formal

```{r}
# Loading required packages
library(purrr)    # For map_df()
library(dplyr)    # For data manipulation
library(timetk)   # For time_series_cv()
library(rsample)  # For analysis() and assessment()
library(here)
library(modelsummary)
library(tidyr)
library(ranger) 
library(knitr)
library(kableExtra)

@@

modelsummary(
  model,
  output = here("results", "tables", "testresult.png"),
  title = "This table represents regresssion results from the test data",
  stars = c('*' = 0.1, '**' = 0.05, '***' = 0.01),
  notes = list(
    "Standard errors are shown in parentheses.",
    "Significance levels: * p < 0.1, ** p < 0.05, *** p < 0.01."
  )
)

@@

    data.frame(
      Train_Years = paste(min(train_cv$Year), max(train_cv$Year), sep = "-"),
      Test_Year = max(test_cv$Year),
      RMSE = rmse_val,
      stringsAsFactors = FALSE
    )
  })
}

@@

```{r}
#view summary of CV results
cv_summary <- results %>%
  summarise(
    Mean_RMSE = mean(RMSE, na.rm = TRUE),
    SD_RMSE = sd(RMSE, na.rm = TRUE),
  )

@@

# Extract variable importance
importance_df <- data.frame(
  Variable = names(rf_full_model$variable.importance),
  Importance = rf_full_model$variable.importance
) %>% 
  arrange(desc(Importance))

# Create importance plot
library(ggplot2)
importance_plot <- ggplot(importance_df, aes(x = reorder(Variable, Importance), 
                         y = Importance)) +  # Fixed this line - removed extra parenthesis
  geom_col(fill = "steelblue") +  # Simplified this line
  coord_flip() +
  labs(title = "Random Forest Variable Importance",
       x = "Predictor Variables",
       y = "Importance (Permutation)") +
  theme_minimal()

# Save the plot
ggsave(here("results", "figures", "rf_importance_plot.png"), 
       plot = importance_plot,
       width = 8, height = 6, dpi = 300)
```

@@

```{r}
# Select only the desired columns
rf_results_table <- rf_results %>%
  select(Train_Years, Test_Year, RMSE, Predictors_Used, mtry_Used)

# Create and save the HTML table
rf_results_html <- kable(rf_results_table, 
                         format = "html",
                         digits = 2,
                         caption = "") %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = FALSE,
                position = "center") %>%
  add_header_above(c("Training Period" = 1, "Test Year" = 1, 
                     "Performance" = 1, "Model Parameters" = 2))

>>>>

```{r}
# Loading required packages
library(purrr)    # For map_df()
library(dplyr)    # For data manipulation
library(timetk)   # For time_series_cv()
library(rsample)  # For analysis() and assessmen()
library(here)
library(ggplot2)
library(modelsummary)
library(tidyr)
library(ranger) 
library(knitr)
library(kableExtra)

@@

# Define a named vector for clean variable names
clean_names <- c(
  "(Intercept)" = "Intercept",
  "GDP_growth_annual_" = "GDP Growth (annual %)",
  "Income_share_held_by_lowest_20" = "Income Share (Lowest 20%)",
  "Control_of_Corruption_Estimate" = "Control of Corruption",
  "Life_expectancy_at_birth_total_years" = "Life Expectancy (years)",
  "Unemployment_youth_total__of_total_labor_force_ages_1524_modeled_ILO_estimate" = "Youth Unemployment (% of labor force, ages 15-24)",
  "Trade__of_GDP" = "Trade (% of GDP)",
  "Foreign_direct_investment_net_inflows_BoP_current_US" = "FDI Net Inflows (current US$)",
  "Access_to_electricity_rural__of_rural_population" = "Rural Electricity Access (% of rural population)",
  "Age_dependency_ratio__of_workingage_population" = "Age Dependency Ratio"
)

# Generate the table with formatted names
modelsummary(
  model,
  output = here("results", "tables", "testresult.png"),
  title = "",
  stars = c('*' = 0.1, '**' = 0.05, '***' = 0.01),
  coef_rename = clean_names,
  notes = list(
    "Standard errors are shown in parentheses.",
    "Significance levels: * p < 0.1, ** p < 0.05, *** p < 0.01."
  )
)

@@

results <- data.frame(
  `Train Years` = paste(min(train_cv$Year), max(train_cv$Year), sep = "-"),
  `Test Year` = max(test_cv$Year),
  RMSE = rmse_val,
  stringsAsFactors = FALSE,
  check.names = FALSE  # Preserves the spaces in column names
)
  })
}

@@

```{r}
#view summary of CV results
cv_summary <- results %>%
  summarise(
    `Mean RMSE` = mean(RMSE, na.rm = TRUE),
    `Std.D (RMSE)` = sd(RMSE, na.rm = TRUE)
  )

@@

# First create a named vector for pretty variable names
neat_names <- c(
  "GDP_growth_annual_" = "GDP Growth",
  "Income_share_held_by_lowest_20" = "Income Share (Lowest 20%)",
  "Control_of_Corruption_Estimate" = "Control of Corruption",
  "Life_expectancy_at_birth_total_years" = "Life Expectancy",
  "Unemployment_youth_total__of_total_labor_force_ages_1524_modeled_ILO_estimate" = "Youth Unemployment",
  "Trade__of_GDP" = "Trade (% of GDP)",
  "Foreign_direct_investment_net_inflows_BoP_current_US" = "FDI Net Inflows",
  "Access_to_electricity_rural__of_rural_population" = "Rural Electricity Access",
  "Age_dependency_ratio__of_workingage_population" = "Age Dependency Ratio"
)

# Extract and prepare importance data with neat names
importance_df <- data.frame(
  Variable = names(rf_full_model$variable.importance),
  Importance = rf_full_model$variable.importance,
  stringsAsFactors = FALSE
) %>% 
  mutate(Variable = neat_names[Variable]) %>%  # Replace with neat names
  arrange(desc(Importance))

# Create importance plot with better formatting
importance_plot <- ggplot(importance_df, aes(x = reorder(Variable, Importance), 
                                           y = Importance)) +
  geom_col(fill = "steelblue", width = 0.7) +
  coord_flip() +
  labs(title = "Random Forest Variable Importance",
       x = NULL,  #  redundant axis title
       y = "Importance (Permutation)") +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    plot.title = element_text(hjust = 0.5, face = "bold")
  )

# Saving the plot
ggsave(here("results", "figures", "rf_importance_plot.png"), 
       plot = importance_plot,
       width = 10,  #  wider for better text fit
       height = 6, 
       dpi = 300)
```

@@

```{r}
# Select and rename columns with clean names
rf_results_table <- rf_results %>%
  select(
    `Train Years` = Train_Years,
    `Test Year` = Test_Year,
    RMSE,
    `Predictors Used` = Predictors_Used,
    `mtry Used` = mtry_Used
  )

# Create and save the HTML table with formatted headers
rf_results_html <- kable(rf_results_table, 
                        format = "html",
                        digits = 2,
                        col.names = c("Train Years", "Test Year", "RMSE", 
                                     "No. of predictors", "mtry"),
                        caption = "") %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = FALSE,
                position = "center") %>%
  add_header_above(c(" " = 1, " " = 1,  # Empty spaces for first two columns
                    "Performance" = 1, 
                    "Model Parameters" = 2)) %>%
  row_spec(0, bold = TRUE)  # Make header row bold

### 2025-05-03 products/manuscript/Manuscript.qmd
made tables formal

```{r}
#| label: tbl-rf-importance-results
#| tbl-cap: "RF importance plot"
#| echo: FALSE
#| out.width: "100%"
knitr::include_graphics(here("results", "figures", "rf_importance_plot.png"))
```

@@

```{r}
#| label: tbl-test-regression-results
#| tbl-cap: "Regression Results for the test dataset"
#| echo: FALSE
#| out.width: "70%"
knitr::include_graphics(here("results", "tables", "testresult.png"))
```

>>>>

```{r}
#| label: tbl-rf-importance-results
#| tbl-cap: "Random Forest importance plot"
#| echo: FALSE
#| out.width: "100%"
knitr::include_graphics(here("results", "figures", "rf_importance_plot.png"))
```

@@

```{r}
#| label: tbl-test-regression-results
#| tbl-cap: "Regression Results for the test dataset"
#| echo: FALSE
#| out.width: "55%"
knitr::include_graphics(here("results", "tables", "testresult.png"))
```
